Extended Data

### 2022-03-21 doc/manuscript/Dow_manuscript_Nature.Rmd
addressing R1 comments
mostly fixing references

@camerondow35 , the quilt plot methods paragraph needs some attention from you

Most studies on tree growth responses to warmer spring temperatures have focused on boreal or temperate conifers, which tend to respond to warmer spring temperatures with an earlier start to growth[@huang_photoperiod_2020;<!--@delpierre_chilling_2019;-->@rossi_critical_2008] and increased annual growth in mesic climates[@cook_predicting_1991;@babst_twentieth_2019;@anderson-teixeira_joint_2021; @gao_earlier_2022]. <!--we can potentially cut some of these references, if not needed elsewhere.-->
In contrast, we have little evidence as to how stem growth and $ANPP_{woody}$ respond to warmer spring temperatures in deciduous forests[@savage_coordination_2021;@delpierre_temperate_2016].
Close coordination of budburst and initiation of xylogenesis[@delpierre_temperate_2016] suggests that warm springs should shift the onset of growth earlier alongside observed advances in leaf phenology[@jeong_phenology_2011;@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020].<!--Most of htese citations are important in other spots, but do we need all of them here?-->
However, earlier initiation of growth would not necessarily translate to earlier, faster, or greater stem growth; rather, stem growth is dependent upon environmental conditions on hourly to daily time scales[@zweifel_why_2021;@tumajer_limitation_2022], and annual growth is more closely linked to conditions within the growing season than to growing season length[@etzold_number_2022], gross primary productivity (GPP)[@delpierre_wood_2016], or NEE[@delpierre_wood_2016]. 
Furthermore, growth of broadleaf deciduous trees may be sink-saturated[@fatichi_moving_2014], such that longer growing seasons with more C fixation do not necessarily augment growth[@delpierre_wood_2016;@zani_increased_2020;@zohner_how_2021].
Tree-ring records, which can be used to examine relationships of annual growth to temperature but not to understand growth seasonality, reveal that growth of temperate deciduous trees tends to be most sensitive to temperature or potential evapotranspiration between late spring and early summer[@dorangeville_drought_2018;@helcoski_growing_2019], with some evidence that warmer springs may have a modest, but usually insignificant, positive effect on growth[@helcoski_growing_2019;@gao_earlier_2022].
These observations do not necessarily align with the finding that warming spring temperatures increase annual forest CO~2~ uptake in temperate deciduous forests[@keenan_net_2014;@lu_no_2022].
Characterizing responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms.

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = `r n_dendro_trees_SCBI` trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = `r n_dendro_trees_HF` trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year (DOY) when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum growth rate ($g_{max}$; occurs on average within 5 days of $DOY_{50}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- <!--(67 trees at SCBI and 138 trees at Harvard Forest)-->and diffuse porous <!--(52 trees at SCBI and 183 trees at Harvard Forest)-->species (`r EDT_species`), which differ in the seasonal timing of growth(`r  EDT_phenology_parameters`)[@dorangeville_peak_2021;@tumajer_limitation_2022]. 
These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).

@@

<!--Seasonal milestones for both canopy foliage phenology and stem growth occurred 3-16 days earlier at SCBI than at Harvard Forest (Fig. `r DOY_timing`, `r  EDT_phenology_parameters`).-->
<!--Consistent with the results of Ref[@dorangeville_peak_2021], ring porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 155-128` days at SCBI and `r 166-144` at Harvard Forest), and their growth was spread over a longer growing season (average $L_{PGS}$ 22 and 17 days longer at SCBI and Harvard Forest, respectively; Fig. `r DOY_timing`, `r  EDF_growth_curves_all`, `r  EDT_phenology_parameters`).--> 
<!--Peak growing season length was similar across sites, with $L_{PGS}$ being, on average, only six days longer at SCBI for ring porous species and less than one day longer for diffuse porous species (`r  EDT_phenology_parameters`).-->

@@

<!--.short vs long-term effects-->
It is possible that as spring warming continues, forests will adjust to directional changes in growing season length with an enhancement of $ANPP_{woody}$.
Across latitudinal gradients, warmer spring temperatures are associated with earlier leaf-out[@delgado_differences_2020] and longer growing seasons, which in turn are correlated with greater tree growth[@anderson-teixeira_joint_2021], woody productivity[@banburymorgan_global_2021], and NEE[@churkina_spatial_2005].
Thus, warming spring temperatures are likely to increase the biophysical potential for annual tree growth.
If extra photosynthate made available through a growing difference between GPP and $ANPP_{woody}$ is allocated to functions that relieve limitations on woody growth -- for example, by enhancing nutrient and water acquisition through enhanced allocation to roots[@zohner_how_2021;@liu_phenological_2022] -- it is possible that warming spring temperatures could ultimately increase $ANPP_{woody}$ through indirect mechanisms.
Understanding how warming spring temperatures are influencing C allocation within ecosystems remains an important outstanding question.

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results clearly demonstrate that the dominant effects of temperature on deciduous tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2021;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), our tree-ring analysis does reveal a higher frequency of negative than positive correlations of annual growth to spring temperatures, particularly for ring porous species in cooler climates (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
Thus, warm springs can have a net negative effect on growth, particularly when water is limiting[@gao_earlier_2022].

@@

Dendrometer band measurements were collected at SCBI[@bourg_initial_2013] and Harvard Forest[@finzi_carbon_2020;@dorangeville_peak_2021], both part of the Forest Global Earth Observatory (ForestGEO)[@anderson-teixeira_ctfsforestgeo_2015;@davies_forestgeo_2021]. 
SCBI (38.8935° N, 78.1454° W; elevation 273–338 m.a.s.l.) is located in the Blue Ridge Mountains at the northern end of Shenandoah National Park, 5 km south of Front Royal, Virginia. 
The forest is secondary and mixed age, having established in the mid-19th century after conversion from agricultural fields[@bourg_initial_2013]. 
Dominant canopy species within the 25.6 ha ForestGEO plot include tulip poplar (*Liriodendron tulipifera* L.), oaks (*Quercus spp.*), and hickories (*Carya spp.*)[@helcoski_growing_2019]. 
The climate is humid temperate, with 1950-2019 mean annual precipitation of 1018 mm and temperatures averaging 1°C in January and 24°C in July[@anderson-teixeira_joint_2021].
Within the study period for the dendrometer band analysis (2011-2020), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-2010). Prior to the study period, the average spring $T_{max}$ was 16.9 $degreeC$ and summer SPEI was -0.06. In the study period, average spring $T_{max}$ was 17.6 $degreeC$ and summer SPEI was 0.51.

Harvard Forest (42.5388° N, 72.1755° W, 340-368 m.a.s.l.) is located near the central Massachusetts town of Petersham. 
The forest is secondary and mixed age, having re-established around the beginning of the 20th century following agricultural use and significant hurricane damage in 1938. 
Dominant species within the 35 ha ForestGEO plot are hemlock (*Tsuga canadensis* (L.) Carrière), oak (*Quercus spp.*) and red maple (*Acer rubrum* L.). 
The climate is temperate continental, with 1950-2019 mean annual precipitation of 1104 mm and temperatures averaging -5° C in January and 22° C in July[@anderson-teixeira_joint_2021]. 
Within the study period for the dendrometer band analysis (1999-2003), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-1998). Prior to the study period, the average spring $T_{max}$ was 10.9 $degreeC$ and summer SPEI was 0.13. In the study period, average spring $T_{max}$ was 11.2 $degreeC$ and summer SPEI was 0.20.
The summer of 1999 was abnormally dry, with precipitation of only 52 mm mo^-1^ in June-August compared to average monthly precipitation of 100+ mm[@dorangeville_peak_2021].

@@

<!--Hypothesis testing -->
Correlation between the dendrometer band-derived growth parameters ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, $L_{PGS}$, $g_{max}$, and $\Delta DBH$, Fig. `r schematic_summary`) and CTW $T_{max}$ (at weekly or monthly resolution, as described above) were assessed using a linear mixed model in a hierarchical Bayesian framework.
Analyses were run for both $T_{max}$ and $T_{min}$, with qualitatively similar results, but we present only results for $T_{max}$, which had overall stronger correlation with growth parameters. 
Mixed effects models were used to test the response of growth parameters to fixed effects of xylem porosity and mean $T_{max}$ (or $T_{min}$) during the CTW, along with random effects of species and of individual tree. 
We ran separate models for each site, and for the response of all growth parameters to $T_{max}$ (or $T_{min}$). 
This mixed-effect model was run within a hierarchical Bayesian framework and fit using the `rstanarm` R interface to the Stan programming language[@gabry_rstanarm_2020;@stan_development_team_stan_2019]. 
In all cases unless otherwise specified, all prior distributions are set to be the weakly informative defaults. 
<!--
More information on rstanarm's default weakly informative priors:
https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html#default-weakly-informative-prior-distributions
-->

@@

<!--QUILT PLOT METHODS-->
For the analysis of correlation between RWI and climate variables, we obtained monthly $T_{max}$ and $T_{min}$ data for 1901-2019 from CRU v.4.04 [@harris_version_2020].
Correlations between monthly climate and RWI were assessed using *dplR*[@bunn_dendrochronology_2008] and *bootRes*[@zang_dendroclimatic_2013] in R v 4.0 (R Core Team, 2020), which correlated functions and bootstrapped confidence intervals for these relationships[@biondi_dendroclim2002_2004].
We analyzed these correlations for January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects of spring temperatures on growth the following year, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).
Correlations and significance levels for months March-August are given in `r chron_table`.

>>>>

Most studies on tree growth responses to warmer spring temperatures have focused on boreal or temperate conifers, which tend to respond to warmer spring temperatures with an earlier start to growth[@huang_photoperiod_2020;<!--@delpierre_chilling_2019;-->@rossi_critical_2008] and increased annual growth in mesic climates[@cook_predicting_1991;@babst_twentieth_2019;@anderson-teixeira_joint_2022; @gao_earlier_2022]. <!--we can potentially cut some of these references, if not needed elsewhere.-->
In contrast, we have little evidence as to how stem growth and $ANPP_{woody}$ respond to warmer spring temperatures in deciduous forests[@savage_coordination_2021;@delpierre_temperate_2016].
Close coordination of budburst and initiation of xylogenesis[@delpierre_temperate_2016] suggests that warm springs should shift the onset of growth earlier alongside observed advances in leaf phenology[@jeong_phenology_2011;@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020].<!--Most of htese citations are important in other spots, but do we need all of them here?-->
However, earlier initiation of growth would not necessarily translate to earlier, faster, or greater stem growth; rather, stem growth is dependent upon environmental conditions on hourly to daily time scales[@zweifel_why_2021;@tumajer_limitation_2022], and annual growth is more closely linked to conditions within the growing season than to growing season length[@etzold_number_2022], gross primary productivity (GPP)[@delpierre_wood_2016], or NEE[@delpierre_wood_2016]. 
Furthermore, growth of broadleaf deciduous trees may be sink-saturated[@fatichi_moving_2014], such that longer growing seasons with more C fixation do not necessarily augment growth[@delpierre_wood_2016;@zani_increased_2020;@zohner_how_2021].
Tree-ring records, which can be used to examine relationships of annual growth to temperature but not to understand growth seasonality, reveal that growth of temperate deciduous trees tends to be most sensitive to temperature or potential evapotranspiration between late spring and early summer[@dorangeville_drought_2018;@helcoski_growing_2019], with some evidence that warmer springs may have a modest, but usually insignificant, positive effect on growth[@helcoski_growing_2019;@gao_earlier_2022].
These observations do not necessarily align with the finding that warming spring temperatures increase annual forest CO~2~ uptake in temperate deciduous forests[@keenan_net_2014;@lu_no_2022].
Characterizing responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms.

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = `r n_dendro_trees_SCBI` trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = `r n_dendro_trees_HF` trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year (DOY) when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum growth rate ($g_{max}$; occurs on average within 5 days of $DOY_{50}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- <!--(67 trees at SCBI and 138 trees at Harvard Forest)-->and diffuse porous <!--(52 trees at SCBI and 183 trees at Harvard Forest)-->species (`r EDT_species`), which differ in the seasonal timing of growth(`r  EDT_phenology_parameters`)[@dorangeville_peak_2022;@tumajer_limitation_2022]. 
These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).

@@

<!--Seasonal milestones for both canopy foliage phenology and stem growth occurred 3-16 days earlier at SCBI than at Harvard Forest (Fig. `r DOY_timing`, `r  EDT_phenology_parameters`).-->
<!--Consistent with the results of Ref[@dorangeville_peak_2022], ring porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 155-128` days at SCBI and `r 166-144` at Harvard Forest), and their growth was spread over a longer growing season (average $L_{PGS}$ 22 and 17 days longer at SCBI and Harvard Forest, respectively; Fig. `r DOY_timing`, `r  EDF_growth_curves_all`, `r  EDT_phenology_parameters`).--> 
<!--Peak growing season length was similar across sites, with $L_{PGS}$ being, on average, only six days longer at SCBI for ring porous species and less than one day longer for diffuse porous species (`r  EDT_phenology_parameters`).-->

@@

<!--.short vs long-term effects-->
It is possible that as spring warming continues, forests will adjust to directional changes in growing season length with an enhancement of $ANPP_{woody}$.
Across latitudinal gradients, warmer spring temperatures are associated with earlier leaf-out[@delgado_differences_2020] and longer growing seasons, which in turn are correlated with greater tree growth[@anderson-teixeira_joint_2022], woody productivity[@banburymorgan_global_2021], and NEE[@churkina_spatial_2005].
Thus, warming spring temperatures are likely to increase the biophysical potential for annual tree growth.
If extra photosynthate made available through a growing difference between GPP and $ANPP_{woody}$ is allocated to functions that relieve limitations on woody growth -- for example, by enhancing nutrient and water acquisition through enhanced allocation to roots[@zohner_how_2021;@liu_phenological_2022] -- it is possible that warming spring temperatures could ultimately increase $ANPP_{woody}$ through indirect mechanisms.
Understanding how warming spring temperatures are influencing C allocation within ecosystems remains an important outstanding question.

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results clearly demonstrate that the dominant effects of temperature on deciduous tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2022;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), our tree-ring analysis does reveal a higher frequency of negative than positive correlations of annual growth to spring temperatures, particularly for ring porous species in cooler climates (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
Thus, warm springs can have a net negative effect on growth, particularly when water is limiting[@gao_earlier_2022].

@@

Dendrometer band measurements were collected at SCBI[@bourg_initial_2013] and Harvard Forest[@finzi_carbon_2020;@dorangeville_peak_2022], both part of the Forest Global Earth Observatory (ForestGEO)[@anderson-teixeira_ctfsforestgeo_2015;@davies_forestgeo_2021]. 
SCBI (38.8935° N, 78.1454° W; elevation 273–338 m.a.s.l.) is located in the Blue Ridge Mountains at the northern end of Shenandoah National Park, 5 km south of Front Royal, Virginia. 
The forest is secondary and mixed age, having established in the mid-19th century after conversion from agricultural fields[@bourg_initial_2013]. 
Dominant canopy species within the 25.6 ha ForestGEO plot include tulip poplar (*Liriodendron tulipifera* L.), oaks (*Quercus spp.*), and hickories (*Carya spp.*)[@helcoski_growing_2019]. 
The climate is humid temperate, with 1950-2019 mean annual precipitation of 1018 mm and temperatures averaging 1°C in January and 24°C in July[@anderson-teixeira_joint_2022].
Within the study period for the dendrometer band analysis (2011-2020), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-2010). Prior to the study period, the average spring $T_{max}$ was 16.9 $degreeC$ and summer SPEI was -0.06. In the study period, average spring $T_{max}$ was 17.6 $degreeC$ and summer SPEI was 0.51.

Harvard Forest (42.5388° N, 72.1755° W, 340-368 m.a.s.l.) is located near the central Massachusetts town of Petersham. 
The forest is secondary and mixed age, having re-established around the beginning of the 20th century following agricultural use and significant hurricane damage in 1938. 
Dominant species within the 35 ha ForestGEO plot are hemlock (*Tsuga canadensis* (L.) Carrière), oak (*Quercus spp.*) and red maple (*Acer rubrum* L.). 
The climate is temperate continental, with 1950-2019 mean annual precipitation of 1104 mm and temperatures averaging -5° C in January and 22° C in July[@anderson-teixeira_joint_2022]. 
Within the study period for the dendrometer band analysis (1999-2003), Spring temperatures (March + April $T_{max}$) and summer SPEI values (4-month value of August) were similar to the average prior to the study period (1970-1998). Prior to the study period, the average spring $T_{max}$ was 10.9 $degreeC$ and summer SPEI was 0.13. In the study period, average spring $T_{max}$ was 11.2 $degreeC$ and summer SPEI was 0.20.
The summer of 1999 was abnormally dry, with precipitation of only 52 mm mo^-1^ in June-August compared to average monthly precipitation of 100+ mm[@dorangeville_peak_2022].

@@

<!--Hypothesis testing -->
Correlation between the dendrometer band-derived growth parameters ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, $L_{PGS}$, $g_{max}$, and $\Delta DBH$, Fig. `r schematic_summary`) and CTW $T_{max}$ (at weekly or monthly resolution, as described above) were assessed using a linear mixed model in a hierarchical Bayesian framework.
Analyses were run for both $T_{max}$ and $T_{min}$, with qualitatively similar results, but we present only results for $T_{max}$, which had overall stronger correlation with growth parameters. 
Mixed effects models were used to test the response of growth parameters to fixed effects of xylem porosity and mean $T_{max}$ (or $T_{min}$) during the CTW, along with random effects of species and of individual tree. 
We ran separate models for each site, and for the response of all growth parameters to $T_{max}$ (or $T_{min}$). 
This mixed-effect model was run within a hierarchical Bayesian framework and fit using the `rstanarm` R interface to the Stan programming language[@gabry_rstanarm_2020;@standevelopmentteam_stan_2019]. 
In all cases unless otherwise specified, all prior distributions are set to be the weakly informative defaults. 
<!--
More information on rstanarm's default weakly informative priors:
https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html#default-weakly-informative-prior-distributions
-->

@@

<!--QUILT PLOT METHODS-->
For the analysis of correlation between RWI and climate variables, we obtained monthly $T_{max}$ and $T_{min}$ data for 1901-2019 from CRU v.4.04 [@harris_version_2020].
<!-- this sentence needs some work-->Correlations between monthly climate and RWI were assessed in R **v 4.0**[@rcoreteam_language_2020]<!--check reference year--> using the packages *dplR*[@bunn_dendrochronology_2008] <!-- did we actually use this package?--> and *bootRes*[@zang_dendroclimatic_2013] to assess correlations, provide bootstrapped confidence intervals for these relationships, and create summary figures (Fig. `r tree_rings`, Extended Data Figures 6-7) following the methods of Ref.[@biondi_dendroclim2002_2004]. 
Our analysis focused on assessing correlations of RWI to months spanning from January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects of spring temperatures on growth the following year, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).
Correlations and significance levels for months March-August are given in `r chron_table`.
